Latent Group Dropout for Multilingual and Multidomain Machine Translation

### 2021-11-15 2022/acl/acl_latex.tex
fin de la discussion
modifies figures, related works

% Remove the "review" option to generate the final version.
\usepackage[review]{acl}

@@

\title{Multi-domain, multilingual translation with latent multi-task group dropout}

@@

\section{Experimental settings}
\subsection{Baselines}
\subsubsection{Multi-domain, multilingual translation}
In both cases, we examine the performance of three following MT systems.
\begin{itemize}
	\item Transformer
	\item Adapter based Transformer
	\item \system{LaMGD} Tranformer
\end{itemize}
For the sake of brevity, we put the detailed descriptions of these systems in Appendix \ref{appendix:c}.
\subsection{Datasets}
\subsubsection{Multi-domain translation}
We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks. We evaluate models on widely used multilingual translation datasets.
\begin{itemize}
	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
\end{itemize}
The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
\section{Results and ablation studies}
\subsection{Multi-domain translation}
According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}
\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploit the proximity between domains. The experiment use the same data as the previous experiment. However, the domain \domain{LAW} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}. If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}. \citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters. We report below the gap between two settings when using \system{LaMGD} Tranformer. 
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{2}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
    \system{Adapter}   \hfill{\footnotesize[+25m]}  & -0.6 & -0.8 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 0.0 & 0.0 \\ 
    \hline
  \end{tabular}
  \caption{Fuzzy domain separation}
  \label{tab:fuzzy}
\end{table*}

Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domain \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}. Our system does not have this loss between two situation. In section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domain.

\subsection{Multilingual translation}
According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively. Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list. However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} in O2M-diverse. 
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&\SB{34.00}&38.30&8.30&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.20&4.00&\SB{13.30}&21.90&32.20&\SB{34.10}&38.50&8.30&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.10&3.80&12.60&\SB{22.80}&31.80&33.40&38.10&8.10&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.50&9.40&20.00&30.60&37.20&38.20&39.30&19.00&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual translation}
  \label{tab:multilingual}
\end{table*}

\subsection{Similarity between dropping masks}
\label{ssec:abalation}
In this section, we compare the sub-networks of the domains or of the language pairs by computing the average of the similarities between the corresponding dropping mask in each layer.

For the multi-domain experiment, we analyze the case of fuzzy domain separation in Section \ref{ssec:fuzzy}. Table \ref{tab:fuzzy-sim} shows that the sub-networks of \domain{law$_1$} and \domain{law$_2$} are identical, which is one of the goals of our method.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{1cm}|*{7}{r|}} \hline
    & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{bank}}& \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} & \multicolumn{1}{c|}{\domain{rel}} &\multicolumn{1}{c|}{\domain{talk}} \\ \hline 
    \domain{med} &1.0&0.80&0.85&0.83&0.83&0.78&0.89\\
    \domain{bank} &0.80&1.0&0.80&0.90&0.90&0.73&0.82\\
    \domain{it} &0.85&0.80&1.0&0.80&0.80&0.78&0.85\\
    \domain{law$_1$} &0.83&0.90&0.80&1.0&\SB{1.0}&0.75&0.82\\ 
    \domain{law$_2$} &0.83&0.90&0.80&\SB{1.0}&1.0&0.75&0.82\\ 
    \domain{rel} &0.78&0.73&0.78&0.75&0.75&1.0&0.77\\
    \domain{talk} &0.89&0.82&0.85&0.82&0.82&0.77&1.0\\
    \hline
  \end{tabular}
  \caption{Multi-domain sub-networks' similarities}
  \label{tab:fuzzy-sim}
\end{table*}

@@

We also plot the languages based on their dropouts in \ref{fig:visual}
\begin{figure*}[htbp]
\begin{subfigure}{.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.97\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.97\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

\section{Related work}
The most adopted paradigm is to have two groups of parameters including task-agnostic parameters and task-specific parameters. Domain adapters and language adapters are strong baselines of this paradigm. However, adapters does not take into account the proximity of the task in their designed architectures.

\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity based parameter sharing. The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.

\textbf{Sparse Transformer}\citet{xian20deep} proposed using different depths for different tasks. The authors aimed to match the depth of the sub-network to the complexity of the task. \citet{Gong21pay,Gong21adaptive} also take interest in the adaptive sparse Transformers which differ the selection of heads in multi-head attention, of layers and of blocks in feed-forward matrices between different tasks. Mixture-of-experts (MoE) is another effective approach of sparse modeling. Within the transformer model, GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}. \citet{william21switch}

\textbf{Adapters} have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}. The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks. Adapters can also be trained unsurpervisedly in multilingual translation \citep{Philip20monolingual}. However, hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}. In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions and erroneous domain tags. Recently, several works have been dedicated to mitigate this weakness through mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.

Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer. This method is very related to the FusionAdapters of \cite{pfeiffer21adapterfusion}. Both approaches assume the final representation is a combination or a choice between shared representation and task-specific representations. The proximity between task is not taken into account in the routing mechanism. Our method proposed a different point of view on the problem of multi-task routing in the underlying network.

@@

Despite several advantages, our method has several weaknesses including the heuristically predefined number of retained groups, $k$ , and the need of domain label in the inference. In future work, we would like to provide an extensive variational framework to model both the number of retained groups, $k$ and the dropping selections. Secondly, to get rid of the requirement of domain information in the inference, we could condition the dropping selection directly on the input $x$ instead of the domain $d$. These two open question are both very interesting and promising.
\bibliography{bibliography}
\bibliographystyle{acl_natbib}
\appendix
\section{Appendix A}
\label{appendix:a}
This section explains how we calculate $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradient $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.

@@

>>>>

% Remove the "review" option to generate the final version.
%\usepackage[review,nohyperref]{acl}
\usepackage[review]{acl}

@@

\title{Latent Group Dropout for Multilingual and Multidomain Machine Translation}

@@

\section{Experimental settings}
\subsection{Baselines}
\subsubsection{Multi-domain, multilingual translation}
In both cases, we examine the performance of three following MT systems.
\begin{itemize}
	\item Transformer
	\item Adapter based Transformer
	\item \system{LaMGD} Tranformer
\end{itemize}
For the sake of brevity, we put the detailed descriptions of these systems in Appendix \ref{appendix:c}.
\subsection{Datasets}
\subsubsection{Multi-domain translation}
We reuse the same data as the recent work of \citet{Pham21revisiting} in multi-domain translation. For the sake of brevity, we leave most of the details of data in Appendix \ref{appendix:c}.
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks. We evaluate models on widely used multilingual translation datasets.
\begin{itemize}
	\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations of \citet{qi18when} between English and eight related languages.
	\item TED8-Diverse. The dataset consists of parallel sentences between English and eight diverse languages as in \citet{Wang20balancing}.
\end{itemize}
We clarify the notation of the languages in multilingual experiments as follows
\begin{itemize}
	\item Diverse: bos (bosnia), mar(marathi),
         hin(hindi), mkd(macedonia), ell(greek),
         bulgaria(bul), french(fra), korean(kor) 
	\item Related: azerbajiani (aze), belarusian(bel),
          glacian(glg), slovak (slk), turkisk (tur),
		  russian(rus), portugese(por),czech(ces)
\end{itemize}
The detailed description of the datasets is provided in Appendix \ref{appendix:c}.
\section{Results and ablation studies}
\subsection{Multi-domain translation}
According to Table \ref{tab:mdmt}, \system{LaMGD} Tranformer shows a significant improvement (+2.78) over the generic system \system{Transformer} with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves equivalent performance in average compared to \system{Adapter}, which is carrefully finetuned with a approximately 12.4M additional parameters.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+12.4m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.28 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.33 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}
\subsubsection{Fuzzy domain separation}
\label{ssec:fuzzy}
In this experiment, we reuse an experiment proposed by \citet{Pham21revisiting} which measures the efficiency of a multi-domain NMT system exploiting the proximity between domains. The experiment uses the same data as the previous experiment. However, the domain \domain{law} is now randomly split into two pseudo domains \domain{law$_1$} and \domain{law$_2$}. If the multi-domain system made full use of the proximity between \domain{law$_1$} and \domain{law$_2$}, there should be no significant difference between the performance of the system trained with six original domains and with seven domains replacing \domain{law} by \domain{law$_1$} and \domain{law$_2$}. \citet{Pham21revisiting} showed a large gap between two training settings when using residual adapters. We report below the gap between two settings when using \system{LaMGD} Tranformer.

\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{2}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} \\ \hline 
    \system{Adapter}   \hfill{\footnotesize[+25m]}  & -0.6 & -0.8 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 0.0 & 0.0 \\ 
    \hline
  \end{tabular}
  \caption{Fuzzy domain separation}
  \label{tab:fuzzy}
\end{table*}

Table \ref{tab:fuzzy} demonstrates a significant loss when training with two pseudo domains \domain{law$_1$} and \domain{law$_2$} compared to training with the real domain \domain{law}. Our system does not have this loss between two situations. In Section \ref{ssec:abalation}, we show that our algorithm outputs the same sub-network for \domain{law$_1$} and \domain{law$_2$}, that allows a full transfer between two domains.

\subsection{Multilingual translation}
According to Table \ref{tab:multilingual}, \system{LaMGD} Tranformer shows a boost of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in O2M-related, M2O-related, M2O-diverse respectively. Significant gains are observed in \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resourced languages in the list. However, \system{LaMGD} Tranformer lags behind multilingual \system{Transformer} and Language Adapters in O2M-diverse. 
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.80 &7.30&20.80&21.10&39.70&19.80&22.60&15.20&18.91 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.30&6.80&21.10&22.00&39.70&20.00&23.00&15.20&19.01 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.20&\SB{9.40}&20.60&\SB{22.80}&39.60&19.60&22.40&15.00&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.40&16.60&28.50&	27.10&43.70&24.60&30.30&25.60&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.10&15.80&28.40&26.80&43.70&24.50&30.20&25.60&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.30&\SB{17.40}&28.60&\SB{28.70}&43.70&24.50&30.70&25.60&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.20&4.0&12.70&22.20&31.80&\SB{34.00}&38.30&8.30&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.20&4.00&\SB{13.30}&21.90&32.20&\SB{34.10}&38.50&8.30&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.10&3.80&12.60&\SB{22.80}&31.80&33.40&38.10&8.10&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.40&9.70&20.50&31.80&37.50&38.70&39.80&19.00&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.50&9.40&20.00&30.60&37.20&38.20&39.30&19.00&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.50}&9.60&\SB{21.50}&32.20&37.70&38.60&40.00&18.90&27.75 \\
    \hline
  \end{tabular}
  \caption{Multilingual translation}
  \label{tab:multilingual}
\end{table*}

\subsection{Similarity between dropping masks}
\label{ssec:abalation}
This section compares the sub-networks of the domains or the language pairs by computing the average of the similarities between the corresponding dropping masks concatenated through all the layers of the underlying model.

For the multi-domain experiment, we analyze the case of fuzzy domain separation in Section \ref{ssec:fuzzy}. Table \ref{tab:fuzzy-sim} shows that the sub-networks of \domain{law$_1$} and \domain{law$_2$} are identical, which is one of the goals of our method. Furthermore, we remark a large distance between \domain{rel} and the rest of the domains, which correlates to the fact that \domain{rel} has the most distant topic to other domains. In Figure ~\ref{fig:domain}, we visualize the domains using their dropping masks concatenated and mapped to 2D space by PCA.

\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{1cm}|*{7}{r|}} \hline
    & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{bank}}& \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{law$_1$}} & \multicolumn{1}{c|}{\domain{law$_2$}} & \multicolumn{1}{c|}{\domain{rel}} &\multicolumn{1}{c|}{\domain{talk}} \\ \hline 
    \domain{med} &1.0&0.80&0.85&0.83&0.83&0.78&0.89\\
    \domain{bank} &0.80&1.0&0.80&0.90&0.90&0.73&0.82\\
    \domain{it} &0.85&0.80&1.0&0.80&0.80&0.78&0.85\\
    \domain{law$_1$} &0.83&0.90&0.80&1.0&\SB{1.0}&0.75&0.82\\ 
    \domain{law$_2$} &0.83&0.90&0.80&\SB{1.0}&1.0&0.75&0.82\\ 
    \domain{rel} &0.78&0.73&0.78&0.75&0.75&1.0&0.77\\
    \domain{talk} &0.89&0.82&0.85&0.82&0.82&0.77&1.0\\
    \hline
  \end{tabular}
  \caption{Multi-domain sub-networks' similarities}
  \label{tab:fuzzy-sim}
\end{table*}

@@

We also plot the languages based on their dropouts in Figure\ref{fig:visual}.
\begin{figure*}[htbp]
\begin{subfigure}{\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.7\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\\
\begin{subfigure}{\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.7\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

\begin{figure*}[htbp]
\includegraphics[width=0.7\textwidth]{multi_domain}
\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:domain}
\end{figure*}

\section{Related work}
Multidomain, multilingual translation

\citep{sen19multilingual,kong21multilingual} used the similar idea of language proximity-based parameter sharing. The authors proposed a multi-decoder model sharing the same encoder among languages and routed languages in different families to different decoders.

\textbf{Sparse Transformer}\citet{xian20deep} proposed using different depths for different tasks. The authors aimed to match the depth of the sub-network to the complexity of the task. \citet{Gong21pay,Gong21adaptive} also take an interest in the adaptive sparse Transformers, which differ the selection of heads in multi-head attention, layers, and blocks in feedforward matrices between different tasks. Mixture-of-experts (MoE) is another effective approach to sparse modeling. Within the transformer model, the GShard model replaces a single feedforward (FFN) sub-layer with an MoE module consisting of multiple FFN sub-layers \citep{lepikhin21gshard}. \citet{william21switch}

\textbf{Adapters} have proven to be very efficient for NLP multi-task modeling \citep{houlsby19parameter,Bapna19simple,Pham20Study,pfeiffer20adapterhub}. The method consists of plugging several adapters to the intermediate layers of a pretrained Transformer and finetuning these adapters on the downstream tasks. Adapters can also be trained unsupervised in multilingual translation \citep{Philip20monolingual}. However, the hard-coded separation between the domains of different tasks leads to a catastrophic forgetting \citep{pfeiffer21adapterfusion}, which is a common problem in multi-task modeling using neural networks \cite{Michael89catastrophic}. In multi-domain translation, recently \citet{Pham21revisiting} demonstrated the brittleness of domain adapters against fuzzy domain separations, out-of-domain distributions, and erroneous domain tags. Recently, several works have aimed to mitigate this weakness through a mixture of expert mechanisms, for example,\cite{pfeiffer21adapterfusion}.

Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer. This method is very related to the FusionAdapters of \cite{pfeiffer21adapterfusion}. Both approaches assume the final representation is a combination or a choice between shared representation and task-specific representations. The proximity between tasks is not taken into account in the routing mechanism. Our method proposed a different point of view on the problem of multi-task routing in the underlying network.

@@

Despite several advantages, our method has several weaknesses including the heuristically predefined number of retained groups, $k$ , and the need of domain label in the inference. In future work, we would like to provide an extensive variational framework to model both the number of retained groups, $k$ and the dropping selections. Secondly, to get rid of the requirement of domain information in the inference, we could condition the dropping selection directly on the input $x$ instead of the domain $d$. These two open questions are both very interesting and promising.
\bibliography{bibliography}
\bibliographystyle{acl_natbib}
\appendix
\section{Appendix A}
\label{appendix:a}
This section explains how we calculate $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradient $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.

@@

\section{\system{LaMGD}'s training algorithm}

\begin{breakablealgorithm}
\caption{Training \system{LaMGD}} \label{alg:mdmt}
\begin{algorithmic}[1]
\REQUIRE {
  \begin{itemize}
    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
  \item[] 
  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
  \item number of retained groups: k, number of groups: $n_p$
 \end{itemize}}
\REPEAT 
\STATE{Pick a batch from domain $d$}
\STATE{Sampling $\forall p,  g_l^d(p) & \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
\IF{$Early\_stopping(Eval\_scores)$}
	\STATE{break}
\ENDIF
\STATE{$i=i+1$}
\UNTIL{$i> iter\_num$}
\end{algorithmic}
\end{breakablealgorithm}
